`State and Parameter Estimation for Spatio-Temporal Neural Fields' or `A Data Driven Framework for Patient-Specific Neural Field Modelling'

### 2010-06-20 ltx/Journal of Neuroengineering/StateAndParameterEstimationNeuralFields_JoNE.tex
edits to paper linking appendices to text

The analytic derivation of the inner product of the \emph{n}-dimensional Gaussians is provided in~\ref{ap:ConvandIn}. Equation~\ref{eq:ReducedForm} can be simplified by exploiting the symmetry (isotropy) of the connectivity kernel where

@@

which is constant $L \times n_{\theta}$ matrix that can be defined analytically, where $n_{\theta}$ is the number of connectivity basis functions. Now substituting equation~\ref{eq:DefPsi} into~\ref{eq:ReducedForm} gives

@@

\section{Discussion}\label{DiscussionSection}
This paper has presented a method for estimating intracortical connectivity and synaptic dynamics from electrophysiological data. The results indicate good estimation performance, where the mean parameter estimates are in good accordance to the actual values. The estimates of the connectivity kernel parameters show relatively broad distributions, which is expected considering the system is driven by an unknown input, modelled as a random disturbance and the observations are corrupted by noise. To overcome the effect of noise, an evoked-response paradigm should be considered when using electrophysiological recordings. Following such a paradigm, data can be epoched into a sufficient number of realisations such that a distribution of the estimates can be formed.

A significant contribution of this paper is the derivation of the basis function decomposition of the neural field. Through the derivation, a finite-dimensional state-space model is formed that describes a continuous neural field. Previous work from Schiff and Sauer~\ref{schiff2008kalman} demonstrated for the first time the applicability of the unscented Kalman filter for state-estimation of a discrete (in space and time) nonlinear spatial-temporal systems. Within their framework, each discrete spatial location forms an element of the state vector to be estimated with the UKF. Therefore, a limitation of this framework lies in an increase in dimensionality with increases in spatial resolution and in the physical size of the neural field. Increases in dimensionality lead to increases in the uncertainty the estimated state, and problems associated with computational complexity. This paper builds on this previous study, where the estimator is constructed to describe a continuous field. The dimensionality of the state vector is independent of the resolution of a spatial discretisation and is governed by the inherent spatial properties of the field. The frequency analysis methods provide a rigorous framework for the placement and the number of the basis functions required for the decomposition. This enables estimation of a larger field, with a lower dimensional state vector. The basis function decomposition also enabled the coupling to be described by an inner product rather than a spatial convolution. This transformation greatly simplifies the computational complexity of the estimator, which is important for the estimator to run in a real-time scenario.

@@

\section{Convolution of two $n$-dimensional Gaussian basis functions}\label{ap:ConvandIn}

@@

\section{Inner Product Bit}

>>>>

The analytic derivation of the inner product of the \emph{n}-dimensional Gaussians is provided in~\ref{ap:InnerProdOfGaussians}. Equation~\ref{eq:ReducedForm} can be simplified by exploiting the symmetry (isotropy) of the connectivity kernel where

@@

which is constant $L \times n_{\theta}$ matrix that can be defined analytically (see~\ref{ap:ConvOfGaussians} for analytic convolution of two Gaussians), where $n_{\theta}$ is the number of connectivity basis functions. Now substituting equation~\ref{eq:DefPsi} into~\ref{eq:ReducedForm} gives

@@

\section{Discussion}\label{DiscussionSection}
This paper has presented a method for estimating intracortical connectivity and synaptic dynamics from electrophysiological data. The results indicate good estimation performance, where the mean parameter estimates are in good accordance to the actual values. The estimates of the connectivity kernel parameters show relatively broad distributions, which is expected considering the system is driven by an unknown input, modelled as a random disturbance and the observations are corrupted by noise. To overcome the effect of noise, an evoked-response paradigm should be considered when using electrophysiological recordings. Following such a paradigm, data can be epoched into a sufficient number of realisations, where each realisation is a response to a stimuli, such that a distribution of the parameter estimates can be formed.

A significant contribution of this paper is the derivation of the basis function decomposition of the neural field model. Through the derivation, a finite-dimensional state-space model is formed that describes a continuous neural field. Previous work from Schiff and Sauer~\ref{schiff2008kalman} demonstrated for the first time the applicability of the unscented Kalman filter for state-estimation of a discrete (in space and time) nonlinear spatial-temporal systems. Within their framework, each discrete spatial location forms an element of the state vector to be estimated with the UKF. Therefore, a limitation of this framework lies in an increase in dimensionality with increases in  the spatial resolution and in the physical size of the neural field. Increases in dimensionality lead to increases in the uncertainty the estimated state, and problems associated with computational complexity. This paper builds on this previous study, where the estimator is constructed to describe a continuous field. The dimensionality of the state vector is independent of the resolution of a spatial discretisation and is governed by the inherent spatial properties of the field. The frequency analysis methods provide a rigorous framework for the placement and the number of the basis functions required for the decomposition. This enables estimation of a larger field, with a lower dimensional state vector. The basis function decomposition also enabled the coupling to be described by an inner product rather than a spatial convolution. This transformation greatly simplifies the computational complexity of the estimator, which is important for the estimator to run in a real-time scenario.

@@

\section{Convolution of two $n$-dimensional Gaussian basis functions}\label{ap:ConvOfGaussians}

@@

\section{Inner Product Bit}\label{ap:InnerProdOfGaussians}
